---
title: "The Magma Blast"
output: 
  flexdashboard::flex_dashboard:
    storyboard: true
    theme: journal
    source_code: embeded
---
```{r setup, include=FALSE}
library(flexdashboard)
```

```{r include=FALSE}
# load packages
library(tidyverse)
library(maps)
library(patchwork)
library(viridis)
library(hrbrthemes)
library(ggplot2)
library(ggthemes)
library(ggridges)
library(tigris)
library(sf)
library(lemon)
library(lubridate)
library(broom)
library(knitr)
library(scales)
library(viridis)
library(shiny)
library(leaflet)
library(DT)
library(plotly)
library(flexdashboard)
library(jsonlite)
library(maptools)
library(ggplot2)
library(dplyr)
library(usmap)
library(scales)
library(glue)
```

```{r include=FALSE}
#Reading in the data
eruptions <- read_csv("../data/eruptions.csv")
events <- read_csv("../data/events.csv")
sulfur <- read_csv("../data/sulfur.csv")
tree_rings <- read_csv("../data/tree_rings.csv")
events <- read_csv("../data/events.csv")
volcano <- read_csv("../data/volcano.csv")
c_and_c <- read.csv("../data/continent_and_countries.csv") 
colnames(c_and_c) <- c("country", "continent")
head(eruptions)
```



### VOLCANOES AROUND THE WORLD



```{r}
template <- "<p>{volcano_name}</p><p>{last_eruption_year}</p"
volcano_html <- volcano %>% 
  mutate(html = glue(template))

volcano_selected <- volcano_html %>% 
  select(volcano_name, latitude, longitude, primary_volcano_type, country, last_eruption_year, html) %>% 
  mutate(popup_info = paste("<b>Volcano name:<b>", volcano_name, "<br/>",
                            "<b>Last eruption year:<b>", last_eruption_year))

total_years <- volcano_selected %>%  
  select(last_eruption_year) %>% 
  mutate(years = str_count(last_eruption_year))
```

```{r the plot}
type <- colorFactor(topo.colors(16), domain = volcano_selected$primary_volcano_type)


leaflet(volcano_selected) %>%  
  addTiles() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addMarkers(lat = ~ latitude,
             lng = ~ longitude,
             clusterOptions = markerClusterOptions()) %>% 
  addCircles(lat = ~ latitude,
             lng = ~ longitude,
             popup = ~ popup_info,
             radius = 5,
             weight = 5,
             color = ~ type(primary_volcano_type)) %>% 
  addLegend(data = volcano_selected,
            title = "Type of volcano",
            pal = type, 
            values = ~ primary_volcano_type,
            opacity = 1) %>% 
  addScaleBar(position = "bottomleft")
volcano_selected <- volcano_html %>% 
  select(volcano_name, latitude, longitude, primary_volcano_type, country, last_eruption_year, html) %>% 
  mutate(popup_info = paste("<b>Volcano name:<b>", volcano_name, "<br/>",
                            "<b>Last eruption year:<b>", last_eruption_year))


total_years <- volcano_selected %>%  
  select(last_eruption_year) %>% 
  mutate(years = str_count(last_eruption_year))


type <- colorFactor(topo.colors(16), domain = volcano_selected$primary_volcano_type)

leaflet(volcano_selected) %>%  
  addTiles() %>% 
  addProviderTiles(providers$CartoDB.Positron) %>% 
  addMarkers(lat = ~ latitude,
             lng = ~ longitude,
             clusterOptions = markerClusterOptions()) %>% 
  addCircles(lat = ~ latitude,
             lng = ~ longitude,
             popup = ~ popup_info,
             radius = 5,
             weight = 5,
             color = ~ type(primary_volcano_type)) %>% 
  addLegend(data = volcano_selected,
            title = "Type of volcano",
            pal = type, 
            values = ~ primary_volcano_type,
            opacity = 1) %>% 
  addScaleBar(position = "bottomleft")
```


---


**Interesting Facts**

Volcanic terrain is the result of the accumulation of erupted lava, unlike mountains 
that are the result of erosion of earth's crust, uplifting, folding, and faulting.
A vent that connects the molten areas or molten rock(magma) are formed through a
series of cracks beneath and within the volcano. Approximately there are 956
potentially active volcanos world wide.
[link](https://www.usgs.gov/programs/VHP/about-volcanoes)

**About the Plot**


* The bubbles are showing the total number of volcanos in those regions

* The color scheme is showing the different types of volcanos

* The popups show the name of the volcano and when it last erupted 


### MAJOR ERUPTIONS SINCE 1700 {data-commentary-width=400}


```{r}
##cleaning eruption table
eruptions_clean <- eruptions %>% 
  filter(eruption_category == "Confirmed Eruption" & start_year >= 1700 & 
           evidence_method_dating %in% c("Historical Observations", "Seismicity")) %>% 
  select("volcano_number", "volcano_name", "eruption_number","eruption_category",
         "start_year", "end_year", "evidence_method_dating", "vei") %>% 
  drop_na() %>% group_by(volcano_name) %>% 
  mutate(number_of_events = n(), 
         max_vei = max(vei)) %>% 
  ungroup()
##cleaning volcano table
volcano_clean <- volcano %>% 
  select("population_within_100_km", "longitude", "latitude", 
         "country","region", "volcano_number", "volcano_name", 
         "primary_volcano_type")
volcano_continent <- volcano_clean %>% 
  full_join(c_and_c, by = "country") %>% 
  group_by(continent) %>% 
  summarise(count = n())
##Joining tables based on cleaned eruption table
final_table <- eruptions_clean %>% 
  left_join(volcano_clean, by = "volcano_name") %>% 
  full_join(c_and_c, by = "country") %>% 
  mutate(volcano_name = fct_reorder(volcano_name, population_within_100_km),
         population = log10(population_within_100_km+1),
         long_lab = round(longitude, 0),
         long_lab_2 = ifelse(long_lab>0, paste0(long_lab, "N"), paste0(-long_lab, "S"))) %>% 
  filter(number_of_events > 5 & number_of_events < 10) %>% 
  drop_na()
final_table$primary_volcano_type <- str_replace_all(final_table$primary_volcano_type, 
                                                    c("Lava dome(s)" ="Lava dome",
                                                  "Stratovolcano(es)"= "Stratovolcano"))
long_lab_2 <- final_table$long_lab_2
```

```{r the_plot}
##Point Plot
left <- final_table %>% 
  ggplot()+
  geom_point(aes(x = start_year, y = volcano_name, size = vei, 
                 color = continent), alpha = 0.7, stroke=1.2)+
  scale_size_area(breaks=c(1,2,3,4,5,6), max_size = 15,name = "Volcano \nExplosivity \nIndex")+
  scale_color_brewer(palette="Dark2", name = "Location (Continent)")+
  scale_x_continuous(breaks=seq(1700, 2020, 40)) +
  scale_y_discrete(expand=c(0.025,0.025))+
  labs(x='', y='')+
  theme_get()+
  guides(color = guide_legend(override.aes = list(size = 18)))+
  theme(legend.position = 'bottom', 
        legend.key = element_blank(),
        legend.text = element_text(size=25),
        legend.title = element_text(face="bold", colour = "black"),
        panel.grid.major.y = element_line(colour= 'black'),
        panel.grid.major.x = element_line(colour = "white"),
        panel.border = element_blank(),
        axis.text = element_text(size=30, colour='black'),
        title = element_text(colour='black', size=50, face='bold', hjust=0),
        plot.subtitle = element_text(colour='white', size=30, face='bold', hjust=0),
        plot.background = element_rect(color='white',fill = "white"),
        panel.background = element_rect(fill = "white",
                                        colour = "white",
                                        linewidth = 5, linetype = "solid"),
        plot.margin=unit(c(0.5, 0.2, 0, 0), 'cm'))+
  ggtitle("Major Eruptions since 1700")
```

```{r plot_2}
##Bar Chart
right <- final_table %>% 
  group_by(volcano_name, population, primary_volcano_type, country, long_lab_2) %>% 
  summarise(count= n()) %>% 
  ungroup() %>% 
  ggplot()+
  geom_col(aes(x = volcano_name, y = population, 
               fill = primary_volcano_type))+
  geom_text(aes(x = volcano_name, y = population, label = country), size = 6,
            hjust = -0.2, color = "black")+
  scale_fill_discrete(name = "Volcano Type")+
  theme_get()+
  guides(fill=guide_legend(ncol=3, override.aes = list(size = 18)))+
  theme(legend.position = "bottom",
        legend.text = element_text(size=25),
        legend.key = element_blank(),
        axis.text.y = element_text(size=30, colour='black', hjust=0.5),
        axis.text.x = element_text(size=30, colour='black'),
        axis.line.x = element_line(color='black'),
        plot.background = element_rect(color='white', fill = "white"),
        title = element_text(colour='black', size=50, face='bold', hjust=0),
        panel.border = element_blank(),
        plot.subtitle = element_text(colour='white', size=30, face='bold', hjust=0),
        panel.background = element_rect(fill = "white",
                                    colour = "white",
                                    linewidth = 0.5, linetype = "solid"),
        plot.margin=unit(c(0.5, 1, 0.2, -0.5), 'cm'))+
  coord_flip()+
  labs(x='', y='Population within 100km of Volcano')+
  scale_y_continuous(expand = c(0, 0), breaks=seq(0, 10, 2.5)) +
  scale_x_discrete(expand=c(0.025,0.025), label = long_lab_2) +
  ggtitle("Location, Population and Country")
```

```{r fig.dim = c(50, 25)}
cowplot::plot_grid(left, right, rel_widths = c(1, 0.7), nrow=1)

```



---

**Interesting Facts**

- There are about `r volcano$volcano_name %>% unique() %>% length()` active volcanoes in the world.

- These volcanoes are distributed across the globe with Asia having the highest number with
`r volcano_continent %>% filter(continent == "Asia") %>% pull(count)` and Oceania having the 
lowest number with `r volcano_continent %>% filter(continent == "Oceania") %>% pull(count)`.

- The Volcanic Explosivity Index (VEI) is a scale that describes the size of explosive
volcanic eruptions based on magnitude and intensity ranging from 1 (lowest) to 6 (highest).

- There are also different types of volcanoes as listed:  `r volcano$primary_volcano_type %>% unique()`.

**About Plot**

- This plot shows major eruptions (VEI and year of eruption) from around the world
since the 1700's while showing the continental location and country of each volcano. The plot also
shows the population within 100,000 km of the volcano telling you the number of people that may be affected
if any of these volcanoes erupt and the volcano type.


### VOLCANOES IN USA
```{r include=FALSE}
glimpse(volcano)
```

```{r include=FALSE}
# filter US data only
usa_volc <- volcano %>% 
  filter(country == "United States")%>%
  filter(region %in% c( "Canada and Western USA", "Alaska", "Hawaii and Pacific Ocean"))%>%
  mutate(region = factor(region, levels = c("Canada and Western USA", 
                                                      "Alaska", 
                                                      "Hawaii and Pacific Ocean"), 
                      labels = c("Western USA", "Alaska", "Hawaii")))%>%
  mutate(subregion = factor(subregion, levels = c("USA (Washington)", "Aleutian Islands", "Alaska Peninsula", "Alaska (southwestern)", "USA (Oregon)", "USA (Idaho)", "USA (Utah)", "USA (New Mexico)", "USA (California)",  "Alaska (eastern)", "Hawaiian Islands", "Alaska (western)", "USA (Arizona)", "USA (Nevada)", "USA (Wyoming)" ), 
                      labels = c("Washington", "Alaska", "Alaska", "Alaska", "Oregon", "Idaho", "Utah", "New Mexico", "California",  "Alaska", "Hawaii", "Alaska", "Arizona", "California", "Wyoming" )))

head(usa_volc)
  
```
```{r include=FALSE}
usa_volc[64, 9] = -180.42
usa_volc[14, 9] = -184.089
```

```{r include=FALSE}
#Merge eruptions info into volcano data bu volcano number
usa_ve<- full_join(usa_volc, eruptions, by = "volcano_number")

usa_ve<-usa_ve%>%
  drop_na(region)%>%
  select(-latitude.y, -longitude.y, volcano_name.y)%>%
  mutate(latitude=latitude.x)%>%
  mutate(longitude=longitude.x)%>%
  mutate(volcano_name= volcano_name.x)%>%
  select(volcano_number, volcano_name, primary_volcano_type:subregion, elevation:population_within_100_km, eruption_number:longitude)%>%
  dplyr::mutate(subregion = forcats::fct_lump(subregion, n = 5))%>%
  mutate(subregion = fct_reorder(subregion, elevation))

head(usa_ve)
```

	



```{r include=FALSE}
# load map for US
usa <- map_data("world", region = "usa", class = "sf")


```


```{r include=FALSE}
#Change to an sf object by saying which columns are the coordinates and setting a CRS:
usa_crs<- st_as_sf(usa_ve, coords = c("longitude", "latitude")) %>%
  st_set_crs(4269) #setting CRS
usa_crs %>% slice(1:3)
```


```{r include=FALSE}
# ridgeline plot of elevation across regions
p1 <- ggplot() +
  geom_density_ridges_gradient(
    data = usa_ve,
    na.rm = TRUE,
    aes(x = elevation, y = subregion, fill = stat(x))
  ) + #population_within_100_km last_eruption_year
  scale_fill_viridis(name = "Elevation (m)", option = "A") +
  theme_minimal() +
  theme(
    legend.position = "none", 
    panel.grid = element_blank(),
    axis.title.y = element_blank(),
    axis.text.x = element_blank(),
    axis.title.x = element_blank()
  )
p1
```


```{r include=FALSE}
# map of volcano locations, colour coded by elevation
p2 <- ggplot() +
  geom_polygon(
    data = usa,
    aes(x = long, y = lat, group = group),
    fill = "lightgrey",
    alpha = 0.6
  ) +
  xlim(c(-186, -65)) + ylim(c(18, 72))+
  
  geom_point(
    data = usa_ve,
    aes(x = longitude, y = latitude, colour = elevation),
  ) +
  scale_color_viridis(name = "Elevation(m)", option= "magma", limits = c(-750, 5000)) +

  
  theme_void()+
  theme(legend.position = "right")
  
p2

```

```{r include=FALSE}
# specify layout for patchwork  
layout <- c(
  area(0, 0, 5, 5), 
  area(0, 4, 2.5, 5.5)
)
```


```{r message=FALSE, warning=FALSE}
# put it all together
p2 +
  p1 + 
  plot_layout(design = layout) +  
  plot_annotation(
    title = "Volcanoes in The United States",
    subtitle = "Histograms of volcano elevations by top 5 subregions",
    caption = "Data from The Smithsonian Institution"
  ) 

#ggsave(filename = "US_volcanoes.png")

```


---

Interesting Facts 

The highest elevation volcano is in Washington named Mt. Rainier elevation 4392m

The lowest elevation if a volcano is in California named Salton Buttes elevation -40m

Fun fact about volcano's in the USA is that they are all on the western side of the continent. The last volcano in this data set to errupt was Semisopochnoi in 2020 in Alaska, its elevation is 1221m. 

About Plot

The volcanos are maped by elevation using a function called map-data with points plotted as sf points. scale_fill_viridis was used to create the color with the theme of magma because of volcanos.

the elevations of the top 5 regions was created with geom_density_ridges_gradient which creates histogram like distributions of the volcano heights. The colorscheme was set to match with the points on the map.




### TIME SERIES PLOT

```{r}
#clean and merge data
eruptions <- read_csv("../data/eruptions.csv")
volcano <- read_csv("../data/volcano.csv")

japan_vol <- volcano %>%
  filter(country == "Japan") %>%
  select(volcano_name, last_eruption_year, country, subregion) %>% #volcano_number, 
  filter(last_eruption_year!="Unknown") %>%
  rename(start_year = "last_eruption_year")

japan_erupt <- eruptions %>%
  #filter(volcano_name == "Kilauea") %>%
  select(volcano_name, vei, start_year) %>%
    filter(!is.na(vei)) #eruption_number, 

  
japan_erupt$start_year <- as.character(japan_erupt$start_year)
  
  #unique()

eruptions_merged <- japan_vol %>%
  inner_join(japan_erupt)

eruptions_merged$start_year <- as.numeric(eruptions_merged$start_year)

eruptions_merged <- eruptions_merged %>%
    filter(start_year > 1930)

# Using for loop to create data needed to plot a geom_polygon
volcano_polygon_list = list()
years = unlist(eruptions_merged$start_year)
volcano_ids = unlist(eruptions_merged$volcano_name)
veis = unlist(eruptions_merged$vei)

for(i in 1:length(years)){
  volcano_polygon_df = data.frame(
    x = c(years[i], years[i] + .75, years[i] + 1.0),
    y = c(0, veis[i], 0),
    t = rep(volcano_ids[i], 3)
  )
  
  volcano_polygon_list[[i]] = volcano_polygon_df
}

# Converting into df and adding subregion information
volcano_polygon_df = volcano_polygon_list %>%
  bind_rows() %>%
  left_join(., select(eruptions_merged, volcano_name, subregion),
            by = c("t" = "volcano_name"))
```

```{r fig.dim = c(10, 5)}
#plot
volcano_timeline = ggplot() +
  geom_polygon(data = volcano_polygon_df, aes(x = x, y = y, group = t, fill = subregion),
               alpha = 0.9, colour = "black") +
  geom_segment(aes(y = 0, yend = 0, x = 1939, xend = 2021), 
               size = 1,
               colour = "black",
               arrow = arrow()) +
  scale_x_continuous(limits = c(1939, 2021),
                     expand = c(0.005, 0.005)) +
  scale_y_continuous(limits = c(0, 3),
                     expand = c(0.005, 0.005)) +
  scale_fill_manual(values = rcartocolor::carto_pal(n = 4, name = "Geyser")) +
  guides(fill = guide_legend(nrow = 1,
                             label.position = "bottom", 
                             keywidth = grid::unit(10, "lines"),
                             keyheight = grid::unit(1, "lines"))) +
  labs(x = NULL, y = "Volcanic Explosion Index", fill = NULL,
       title = "A Timeline of Volcanic Activity Since 1940 in Subregions of Japan",
       subtitle = "Each triangle represents an eruption",
       caption = paste0("Source: The Smithsonian Institution\n",
                     "Visualization: Ijeamaka Anyene | @ijeamaka_a")) +
  theme(#plot.background = element_rect(fill = "#303030"),
        plot.title = element_text(colour = "#303030", hjust = 0.5, face = "bold", size = 18),
        plot.subtitle = element_text(colour = "#636363", hjust = 0.5, face = "bold", size = 12),
        plot.caption = element_text(colour = "#303030", size = 6),
        panel.border = element_blank(),
        panel.background = element_blank(),
        panel.grid = element_blank(),
        axis.text = element_text(colour = "#636363", face = "bold", size = 14),
        axis.text.y = element_text(colour = "#636363", face = "bold", size = 10),
        legend.position = "bottom",
        legend.text = element_text(colour = "#636363", size = 12),
        axis.title.y = element_text(colour = "#303030", size = 14, face = "bold"),
        #legend.background = element_rect(fill = "#B3B3B3"),
        )
volcano_timeline



# Ref Code https://github.com/Ijeamakaanyene/tidytuesday/blob/master/scripts/2020_10_volcanos.Rmd
```


---


**Interesting Facts**

Japan is located on the "Ring of Fire" which is a chain of volcanoes that stretches 40,000 kilometers. There are over 111 active volcanoes in Japan. The majority of mountains in Japan were formed as volcanoes.

[link](https://www.japan.travel/national-parks/plan-your-visit/guides-and-stories/volcanoes-what-are-they/)

**About the Plot**

* Each triangle represents an eruption

* The height of each triangle represents the volcanic explosion index

* Each color represents the subregion the eruption occured in